---
title: "Exploring Probability Distributions in R"
output: html_document  # Use `output: quarto::html` for Quarto
---

## Introduction

This document provides an overview of the `d-`, `p-`, `q-`, and `r-` functions in R,
focusing on their roles in understanding probability distributions.

## Theoretical Background

### The `d-` Function (Density)
The `d-` functions, such as `dnorm()`, calculate the density (or probability mass)
at a given point. For continuous distributions, this represents the height
of the probability density function (PDF) at a specific value.
Use it to understand the likelihood of observing a specific value under a given distribution.
For example, dnorm(0, mean = 0, sd = 1) returns the density at x = 0.
It helps assess how probable a particular outcome is.

### The `p-` Function (Cumulative Distribution)
The `p-` functions, like `pnorm()`, return the cumulative probability up
to a given point. This is known as the cumulative distribution function (CDF),
which tells us the probability of a random variable being
less than or equal to a specific value.
Function (Cumulative Probability):
calculates the cumulative probability up to a specific point,
indicating the probability of a value being less than or equal to that point.
For instance, pnorm(1, mean = 0, sd = 1) gives the probability P(X≤1).
It's useful for finding thresholds or quantifying areas under the curve.

### The `q-` Function (Quantile)
The `q-` functions, such as `qnorm()`, provide the inverse of the CDF.
Given a probability, `q-` functions return the quantile that corresponds to that probability.
q- Function (Quantile): calculates the value corresponding to a given
cumulative probability. Use it when you need the value for a specified percentile.
For example, qnorm(0.975, mean = 0, sd = 1) returns the 97.5th percentile
of the standard normal distribution, useful in setting confidence intervals.

### The `r-` Function (Random Generation)
The `r-` functions, like `rnorm()`, are used for generating random samples from a
specified distribution. This is essential for simulations and statistical analysis.

## Visualizing a Standard Normal Distribution

Let's create a histogram of random samples drawn from a standard normal
distribution and overlay it with the PDF, CDF, and empirical density.

```{r setup, include=FALSE}
mean <- 0
sd <- 1
set.seed(123)
```

### Generating Random Samples

```{r random-samples}
samples <- rnorm(1000, mean, sd)
```

### Plotting the Histogram with Overlays

```{r plot-histogram}
library(ggplot2)
x <- seq(-4, 4, length.out = 100)
df <- data.frame(x = x, value = samples, group = "Measured")
# Sequence for plotting PDF and CDF and empirical density
densi_emp <- density(samples)

# Normalize the measured points
hist_data <- hist(samples, breaks = 30, plot = FALSE)
N <- length(samples)
n_counts_in_bins <- hist_data$counts
bin_widths <- hist_data$breaks[2] - hist_data$breaks[1]
densi <- n_counts_in_bins / (bin_widths * N)

df <- rbind(df,
  data.frame(x = x, value = dnorm(x, mean, sd), group = "PDF"),
  data.frame(x = x, value = pnorm(x, mean, sd), group = "CDF"),
  data.frame(x = densi_emp$x, value = densi_emp$y, group = "Empirical Density")
)

ggplot() +
  geom_histogram(
    data = df[df$group == "Measured", ],
    aes(x = value, y = ..density..) 
    # This normalizes the histogram: n/(binwidth*N) = density;
    # Where N is the total number of observations and
    # n is the number of observations in each bin.
    # See above for more details
  ) +
  geom_line(
    data = df[df$group != "Measured", ],
    aes(x = x, y = value, color = group),
    size = 1.5
  ) +
  theme(legend.title = element_blank())
```
