---
format: pdf
header-includes:
  \usepackage{fancyhdr}
  \pagestyle{fancy}
params:
  file: "nofile"
bibliography: references.bib  
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

\fancyhf{}
\fancyhead[L]{Cytotoxicity screen}
\fancyfoot[CE,CO]{\thepage}

# Screening Report COMPOUND PLATFORM

\
\

```{r, results='asis', echo = FALSE}
df <- readRDS(params$file)
cat(df[[2]], "\n")
```
\

**Type: Cytotoxicity screen** \
**Cell viability of HeLa cells after exposition to compounds for three days**

\
\


\tableofcontents 


\newpage

#	Introduction

This report summarizes the result of the cytotoxicity test routinely conducted at our insitut *IBCS-FMS* in the working group *ComPlat* at the *KIT*. \

In order to measure the cytotoxicity for mammalian organisms, *HeLa* cells are used within the assay. First, 100 µL of a solution containing 10^5^ cells/mL are sowed into 96 well plates. Subsequently, after 24 h, different concentrations (0.5 µM, 5µM, 10µM and 25 µM; diluted in cell media) of the respective compounds are added to the cells. After incubation of three days, MTT [3-(4,5-Dimethylthiazol-2-yl)-2,5-Diphenyltetrazoliumbromid)] is added. The cells reduce MTT to a formazan by oxidizing the reductant equivalents NADH or NADPH. Likewise, electrons from succinate are used for the reduction of MTT. This step is catalyzed by the enzyme succinate-dehydrogenase. The reaction is stopped after 3 hours, using an aqueous solution consisting of 10% SDS and roughly 0.3% HCl. The last step is the absorbance measurement of the formazan at 595 nm. \

As a positiv control cells were lysed using a 20% triton solution, directly before adding MTT. Cells incubated exclusively in cell medium served as a negative control. \

The cell viability is calculated using the following equation:

$$
cell\ viability_{well_{i}} = \frac{absorbance_{well_{i}} - mean(positive\ control)}{mean(negative\ control)}
$$
\

# Representation of data analysis

The screening results are shown in the next section. The data is represented in one plots shown at the right column. The normalized data is depicted in the Boxplots which also include the corresponding fit. The graphics are produced by using ggplot2 (for details see @ggplotBook). 
The fit is calculated based on the equation below. For further details please refer to @drc and @Seber1989. As a side note, the outliers were detected using a method described by @motulsky_detecting_2006. 

$$
\sum_{i=1}^N [(y_i-f_i) / w_i]^2
$$
where:

* $y_i$ is the ith observation of the measured data
* $w_i$ is the weight for the ith observation
* $f(i) = c + \frac{d-c}{1+\exp(b(\log(i)-\log(e)))}$ is the expected ith value 

The following parameter are listed in the table beneath the plot:
  
  * The estimated residual standard error (*RSE*)
  * The response at the lowest and highest concentration 
  * The $-log10(IC50/1000000)$ value (*pIC50*) 
  * The p value compares the dose-response model with a simple linear regression model with slope 0 (a horizontal regression line corresponding     to no dose effect)
  * The *IC50_relative_higher/lower* arre the $IC50 \pm 95\%$ confidence intervall
  * The *relative IC50* (*e* in the equation above)
  * The Hill coefficent (*b* in the equation above)
  * The first and second asymptote (parameter *c* and *d* respecitvly in the equation above)
  
\newpage

# Compounds and results


```{r}
#| echo: false
#| warning: false
library(cowplot)
MTT:::report_plots(df[[1]])
```

\newpage

# References
